My title

### 2024-11-25 paper/paper.qmd
update model section and result section

```{r}
# Calculate diagnostic measures
glm_model <- glm
# Assuming `glm_model` is your fitted model

# 1. Linearity: Residuals vs Fitted Plot
plot(glm_model, which = 1, main = "Linearity Check: Residuals vs Fitted")
plot(glm_log, which = 1, main = "Linearity Check: Residuals vs Fitted")

# Look for a random scatter of points. Patterns indicate non-linearity.

# 2. Homoscedasticity: Scale-Location Plot
plot(glm_model, which = 3, main = "Homoscedasticity Check: Scale-Location")
plot(glm_log, which = 3, main = "Homoscedasticity Check: Scale-Location")
# Points should be evenly spread. A funnel shape indicates heteroscedasticity.

# 3. Normality of Residuals: Q-Q Plot
plot(glm_model, which = 2, main = "Normality Check: Q-Q Plot")
plot(glm_log, which = 2, main = "Normality Check: Q-Q Plot")

# Points should lie approximately along the diagonal line.

# 4. Residual Histogram: Another Normality Check
residuals <- residuals(glm_model)
hist(residuals, main = "Histogram of Residuals", xlab = "Residuals")
residuals <- residuals(glm_log)
hist(residuals, main = "Histogram of Residuals", xlab = "Residuals")

# Check for symmetry. A skewed histogram suggests non-normal residuals.

# 5. Multicollinearity: Variance Inflation Factor (VIF)
vif_values <- vif(glm_model)
print(vif_values)
vif_values <- vif(glm_log)
print(vif_values)
# VIF > 5 indicates high multicollinearity. Consider removing highly correlated predictors.

# 6. Cook's Distance: Influence of Observations
cooksd <- cooks.distance(glm_model)
plot(cooksd, main = "Cook's Distance", type = "h")
abline(h = 4 / nrow(model.frame(glm_model)), col = "red", lty = 2)

cooksd <- cooks.distance(glm_log)
plot(cooksd, main = "Cook's Distance", type = "h")
abline(h = 4 / nrow(model.frame(glm_log)), col = "red", lty = 2)
# Points above the red line are highly influential.

# 7. Leverage: Identify High Leverage Points
leverage <- hatvalues(glm_model)
plot(leverage, main = "Leverage Check", type = "h")
abline(h = 2 * mean(leverage), col = "red", lty = 2)

leverage <- hatvalues(glm_log)
plot(leverage, main = "Leverage Check", type = "h")
abline(h = 2 * mean(leverage), col = "red", lty = 2)
# Points above the red line have high leverage.

AIC(glm_model)
AIC(glm_log)
BIC(glm_model)
BIC(glm_log)

@@

The model is mathematically expressed as:

@@

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

@@

@@

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

@@

Weaknesses and next steps should also be included.

@@

# Model Details {#sec-model-detail}
## MLR Model With Every Predictor in Cleaned Data

@@

```{r}
#| label: tbl-brm
#| tbl-cap: Summary of Bayesian Model and Comparison of RMSE / MAE
#| tbl-subcap: ['Bayesian Model Shows Similar R2 value Compare to Linear', 'Linear Model shows better fit']
#| echo: false
#| warning: false
#| layout-ncol: 2
modelsummary(brm)

>>>>

@@

The Model LP is mathematically expressed as:

@@

### Model Validation {#sec-validation}

The linear regression model was chosen to analyze the relationship between wind speed, atmospheric pressure, total precipitation, and gust speed with the log-transformed mean temperature. The primary goal of this model is to quantify the effect of each predictor on the response variable and to make predictions. The linear model is suitable because the relationship between the predictors and the response variable was found to be approximately linear after applying logarithmic and Box-Cox transformations to address non-linearity and skewness.

#### Assumptions Fit the Data

```{r}
#| label: fig-assumption
#| fig-cap: Model LP meets all assumtions of linear regression
#| fig-subcap: ['Linearity Check: Residuals vs Fitted', 'Normality Check: Q-Q Plot', 'Homoscedasticity Check: Scale-Location']
#| echo: false
#| warning: false
#| layout-ncol: 2

plot(glm, which = 1)
plot(glm, which = 2)
plot(glm, which = 3)
```

The model, which uses linear regression, has four assumptions: linearity, homoscedasticity, and approximate normality of residuals.

The linearity assumption states that the relationship between each predictor and the response variable is linear. This assumption was evaluated by examining scatterplots of the predictors against the response variable and by analyzing residuals versus fitted values. As shown in @fig-assumption-1, The red smooth line indicates only a slight curvature, and the the residuals scatter reasonably randomly around the 0 line. The resulting diagnostic plots showed no obvious systematic patterns, indicating that the linearity assumption is reasonably satisfied.

Homoscedasticity requires that the variance of the residuals remains constant across all levels of the predictors. In @fig-assumption-3, the points (standardized residuals) are evenly spread across the range of fitted values. There is no clear fan or funnel shape, suggesting that the variance of the residuals is fairly constant, which supports the assumption of homoscedasticity.

Linear regression assumes that the residuals follow a normal distribution. This assumption was assessed using Q-Q plots in @fig-assumption-2, where the residuals were compared to a theoretical normal distribution. Most points aligned closely with the diagonal line, indicating that the residuals are approximately normally distributed.

Independence assumes that each observation in the dataset is unrelated to others. This was ensured during data collection or preparation according to ECCC's methodology as discussed in @sec-methodology.

Multicollinearity occurs when predictors are highly correlated, making it difficult to isolate their individual effects. Variance Inflation Factor (VIF) quantifies how much multicollinearity inflates the variance of the estimated regression coefficients. The Generalized VIF (GVIF) is used because each polynomial term represents more than one degree of freedom. The result in @tbl-vif are below the commonly accepted threshold of 5, indicating low multicollinearity in the predictors.

```{r}
#| label: tbl-vif
#| tbl-cap: VIF Value Indicates Low Multicollinearity
#| echo: false
#| warning: false

# Calculate VIF
vif_values <- vif(glm)
kable(vif_values)
```

#### Validation on Test Data

We validate our Model LP with test data to evaluate the model's generalizability and performance on unseen data.

The R2 value calculated on the test data ($R^2 = 0.59$) is very close to the R2 value of the model on the training data ($R^2 = 0.61$), as discussed in @sec-result-r2 This similarity indicates that the model generalizes well to unseen data, as its performance on the test set is consistent with its performance during training. The small difference between the two R2 values suggests that the model is not overfitting the training data and captures the underlying relationships in the data effectively.

```{r}
#| label: tbl-test-validate
#| tbl-cap: R2 value on test data is close to train data
#| echo: false
#| warning: false

test_data$predicted <- predict(glm, newdata = test_data)
SSR <- sum((test_data$log_mean_temp - test_data$predicted)^2)  # Sum of Squared Residuals
TSS <- sum((test_data$log_mean_temp - mean(test_data$log_mean_temp))^2)  # Total Sum of Squares
R2 <- 1 - (SSR / TSS)
R2T <- summary(glm)$r.squared
results <- data.frame(
  Metric = c("SSR", "TSS", "R2", "R2-Train"),
  Value = c(SSR, TSS, R2, R2T)
)

kable(results)
```

### Model Justification

The decision to use a linear regression model with polynomial terms for this analysis is grounded in the need to capture non-linear relationships between the predictors and the log-transformed mean temperature (log_mean_temp). Initial exploratory data analysis and diagnostic plots indicated that the relationships between the predictors, such as log-transformed wind speed, pressure, total precipitation, and gust speed, and the response variable were not strictly linear. Applying second-degree polynomial terms allows the model to account for curvature and interactions within the data while maintaining interpretability.

The polynomial linear regression model was selected over alternative non-linear methods due to its balance between complexity and simplicity. Polynomial terms extend the linear model's capacity to capture non-linear patterns while preserving the interpretability of regression coefficients. The inclusion of transformations, such as logarithmic and Box-Cox, further improves the model by addressing issues of skewness and heteroscedasticity observed in the raw data.

Finally, the model assumptions were rigorously evaluated in @sec-validation. Residual plots confirmed that the linearity assumption was largely met, while the Scale-Location plot and diagnostic tests supported the homoscedasticity of residuals. Variance Inflation Factor (VIF) values for the predictors were all well below the critical threshold, indicating no significant multicollinearity. These results ensure that the model remains valid and reliable for inference and prediction.

@@

```{r}
#| label: tbl-modelresults
#| tbl-cap: Summary of Model LP
#| echo: false
#| warning: false
modelsummary(glm)
```

## Intercept Indicates a Strong Baseline In Temperature

The intercept of the model is estimated at **3.899**, representing the predicted value of the log-transformed mean temperature when all predictors are at their mean (after centering and scaling in the polynomial terms). This value serves as the baseline against which the effects of all predictors are measured. A significant intercept indicates the model reliably predicts the baseline log-mean temperature.

## Wind Speed: Non-Linear Increase and Decline in Temperature
The first-degree polynomial term for log-transformed wind speed ($\text{poly(log\_wind\_speed, 2)1}$) has a positive coefficient of **0.260**, suggesting that increasing wind speed is associated with a slight increase in $\log(\text{mean\_temp})$. However, the second-degree term ($\text{poly(log\_wind\_speed, 2)2}$) has a negative coefficient of **-0.449**, indicating a reversing effect at higher wind speeds. This quadratic relationship reflects non-linear relationships between wind speed and temperature.

## Atmospheric Pressure: Consistent Negative Impact on Temperature
Log-transformed atmospheric pressure shows a consistently negative relationship with $\log(\text{mean\_temp})$. The first-degree term ($\text{poly(log\_pressure, 2)1}$) has a coefficient of **-0.388**, and the second-degree term ($\text{poly(log\_pressure, 2)2}$) is more negative at **-0.973**. This indicates that higher atmospheric pressure is associated with a reduction in $\log(\text{mean\_temp})$, and the quadratic term amplifies this effect, particularly at extreme values of pressure.

## Total Precipitation: The Strongest Negative Effect
The Box-Cox transformed total precipitation has the strongest impact on $\log(\text{mean\_temp})$ among all predictors. The first-degree term ($\text{poly(total\_precipitation\_boxcox, 2)1}$) has a significant negative coefficient of **-2.016**, implying that higher precipitation reduces $\log(\text{mean\_temp})$ substantially. The second-degree term ($\text{poly(total\_precipitation\_boxcox, 2)2}$) is positive at **0.218**, indicating a slight mitigation of this negative effect at extreme precipitation levels.

## Gust Speed: Decreases Temperature but Shows Curvature
The first-degree term for log-transformed gust speed ($\text{poly(log\_gust\_speed, 2)1}$) has a negative coefficient of **-0.980**, showing that increasing gust speed reduces $\log(\text{mean\_temp})$. However, the second-degree term ($\text{poly(log\_gust\_speed, 2)2}$) has a positive coefficient of **0.347**, suggesting a minor curvature in the relationship where the negative effect is less pronounced at higher gust speeds.

## Model Metrics: Explains 61% Variance with Low RMSE
The model explains a significant portion of the variance in $\log(\text{mean\_temp})$, as indicated by an $R^2$ value of **0.610** and an adjusted $R^2$ of **0.603**. These values demonstrate that approximately 61% of the variability in $\log(\text{mean\_temp})$ is explained by the predictors, even after accounting for the complexity of the model.

The model has a low RMSE of **0.12**, indicating accurate predictions with minimal average error. The AIC (**-601.2**) and BIC (**-560.6**) values suggest a strong fit compared to alternative models.

@@

## Summary of Results: Precipitation Leads, While Wind and Pressure Add Complexity {#sec-first-point}

This polynomial regression model captures non-linear relationships between predictors and $\log(\text{mean\_temp})$ effectively, with total precipitation exerting the strongest influence. Wind speed and gust speed contribute with both positive and negative effects, depending on the quadratic curvature. Atmospheric pressure consistently shows a negative impact. Despite its strengths, slight evidence of heteroscedasticity and non-linear residual patterns observed in diagnostic plots suggest that future refinements, such as interaction terms or alternative modeling techniques, could improve the model's performance.

## The finding does not agree with Visser
## the finding does match with Wills

@@

### Limitations on the Model LP

While the polynomial linear regression model captures much of the relationship between the predictors and the log-transformed mean temperature, some limitations remain. @fig-assumption-1 indicates that the relationship is not completely linear, with minor deviations suggesting that the model may not fully account for certain non-linearities. Additionally, @fig-assumption-3 shows slight evidence of heteroscedasticity at the extreme fitted values, indicating that the variance of residuals is not entirely constant. These limitations suggest that the model could benefit from further refinement, such as applying additional transformations, exploring interaction terms, or using alternative techniques like weighted least squares or robust regression to address potential heteroscedasticity. 
Weaknesses and next steps should also be included.

### Limitation on the Response Mean Temperature

### Limitation on only researching temperature

@@

# Additional Model Details {#sec-model-detail}
## MLR Model With Every Predictor in Cleaned Data

@@

```{r}
#| label: tbl-brm
#| tbl-cap: Summary of Bayesian Model and Comparison of RMSE / MAE on Test Data
#| tbl-subcap: ['Bayesian Model Shows Similar R2 value Compare to Linear', 'Linear Model shows better fit']
#| echo: false
#| warning: false
#| layout-ncol: 2
modelsummary(brm)
